---
title: "ajw_ma_mag_rel_ab_figures_11_8_23"
author: "A.J Wing"
date: "2023-11-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(data.table)
library(tidyr)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(jcolors)
library(viridis)
library(cowplot)
library(gridExtra)
library(vegan)

```

```{r}
metadata <- read_tsv("~/Desktop/erie_story_r_work/2014_story_metadata_updated_12_15_21.tsv")
names(metadata)[names(metadata) == 'samples'] <- "sample"
metadata$sample <- gsub("Sample_","",metadata$sample)
```

##aug host abundances
```{r}
setwd("~/Desktop/erie_story_r_work/aug_ma_mags_coverm_results/")
```

```{r}
files = list.files(pattern="*_ma_mags_relative_abundance.tsv",recursive=TRUE)

blast_hits <- function(x) {
  # Check if file is empty
  if (file.info(x)$size == 0) {
    message(paste0("Skipping empty file ", x))
    return(NULL)
  } else {
    tmp <- read.delim(x, header = FALSE, sep = "\t")
    tmp$sample <- gsub("", "", gsub("_ma_mags_relative_abundance.tsv", "", basename(x)))
    tmp$sample <- gsub("", "", gsub("sample_", "", basename(x)))
    return(tmp)
  }
}

tables <- Filter(NROW, lapply(files, blast_hits))
coverm_data <- rbindlist(tables)
```


##modify giant coverm file to contain meaningful columns
```{r}

names(coverm_data)[names(coverm_data) == 'V1'] <- "host_genome"
names(coverm_data)[names(coverm_data) == 'V2'] <- "host_rel_ab"
names(coverm_data)[names(coverm_data) == 'V3'] <- "host_mean_read_depth"
names(coverm_data)[names(coverm_data) == 'V4'] <- "host_covered_bases"
names(coverm_data)[names(coverm_data) == 'V5'] <- "host_variance"
names(coverm_data)[names(coverm_data) == 'V6'] <- "host_genome_length"
names(coverm_data)[names(coverm_data) == 'V7'] <- "host_rpkm"
names(coverm_data)[names(coverm_data) == 'V8'] <- "host_tpm"
names(coverm_data)[names(coverm_data) == 'V9'] <- "host_sample"

```

```{r}
coverm_data <- coverm_data[coverm_data$host_genome != 'unmapped',]
coverm_data <- coverm_data[coverm_data$host_genome != 'Genome',]
coverm_data$host_genome <- gsub("concatenated_concatenated_","",coverm_data$host_genome)
coverm_data$sample <- gsub("_ma_mags_relative_abundance.tsv","",coverm_data$sample)

coverm_data <- subset(coverm_data, grepl("contigs", host_genome))
```

##merge metadata and coverm data
```{r}
coverm_aug_mags <- merge(coverm_data, metadata, by = "sample")

coverm_aug_mags$host_genome <- gsub("_ma_contigs", "", coverm_aug_mags$host_genome)

coverm_aug_mags$host_rel_ab <- as.numeric(coverm_aug_mags$host_rel_ab)

```



```{r}
# Define a palette of 10 colors
my_palette <- c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3", "#FF7F00", "#FFFF33", "#A65628", "#F781BF", "#999999", "#000000")
fraction_colors <- c("#29BF12", "#00A5CF", "#DE1A1A", "#FFBF00")

p1 <- ggplot(coverm_aug_mags, aes(x = host_genome, y = host_rel_ab, fill = Station)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ fraction, nrow = 1, scales = "free") +  # Facet by Station
  labs(x = "Host Genome", y = "Relative Abundance") +
  scale_fill_manual(values = fraction_colors) +  # Use the defined palette for coloring Station
  theme_minimal() +
  ggtitle("August 4") +  # Add the plot title
  ylim(0, 40) +  # Set y-axis limits to 0 and 40
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels by 45 degrees

p1
```

```{r}
# Define a palette of 10 colors
my_palette <- c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3", "#FF7F00", "#FFFF33", "#A65628", "#F781BF", "#999999", "#000000")
fraction_colors <- c("#29BF12", "#00A5CF", "#DE1A1A", "#FFBF00")

p2 <- ggplot(coverm_aug_mags, aes(x = host_genome, y = host_rel_ab, fill = "4-Aug-14")) +
  geom_bar(stat = "identity") +
  facet_grid(Station ~ fraction, scales = "free") +  # Facet by both Station and fraction
  labs(x = "Host Genome", y = "Relative Abundance") +
  scale_fill_manual(values = "olivedrab") +  # Use the defined palette for coloring host_genome
  theme_minimal() +
  ggtitle("August 4") +  # Add the plot title
  ylim(0, 40) +  # Set y-axis limits to 0 and 40
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p2
```

```{r}
fraction_colors <- c("#29BF12", "#00A5CF", "#DE1A1A", "#FFBF00")

p2 <- ggplot(coverm_aug_mags, aes(x = factor(1), y = host_rel_ab)) +  # Set x to a constant to center the box plots
  geom_boxplot(aes(fill = fraction), color = "black", width = 0.5) +  # Set the fill color by fraction and line color to black, adjust box width if needed
  geom_jitter(aes(fill = fraction), width = 0.2, alpha = 0.5, shape = 21, color = "black") +  # Add jittered points with fill color by fraction and black outline
  facet_wrap(~ interaction(Station, fraction), scales = "free", strip.position = "bottom") +  # Use facet_wrap and place labels at the bottom
  labs(x = "Station and Fraction", y = "Relative Abundance") +
  scale_fill_manual(values = fraction_colors) +  # Use the defined palette for coloring fractions
  theme_minimal() +
  ggtitle("August 4") +  # Add the plot title
  ylim(0, 30) +  # Set y-axis limits to 0 and 30
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        strip.background = element_blank(),  # Remove background for facet labels
        strip.placement = "outside",  # Place facet labels outside the plot area
        panel.grid = element_blank())  # Remove the grid background

p2

```

```{r}
# Define the colors for each fraction
fraction_colors <- c("100" = "#29BF12", "3" = "#00A5CF", "53" = "#DE1A1A", "Whole" = "#FFBF00")

# Reorder the factors for Station and fraction
coverm_aug_mags$Station <- factor(coverm_aug_mags$Station, levels = c("WLE12", "WLE2", "WLE4"))
coverm_aug_mags$fraction <- factor(coverm_aug_mags$fraction, levels = c("Whole", "3", "53", "100"))

p2 <- ggplot(coverm_aug_mags, aes(x = factor(1), y = host_rel_ab)) +  # Set x to a constant to center the box plots
  geom_boxplot(aes(fill = fraction), color = "black", width = 0.5) +  # Set the fill color by fraction and line color to black, adjust box width if needed
  geom_jitter(aes(fill = fraction), width = 0.2, alpha = 0.5, shape = 21, color = "black") +  # Add jittered points with fill color by fraction and black outline
  facet_grid(Station ~ fraction, scales = "free") +  # Use facet_grid to arrange panels as specified
  labs(x = "Station and Fraction", y = "Relative Abundance") +
  scale_fill_manual(values = fraction_colors) +  # Use the named palette for coloring fractions
  theme_minimal() +
  ggtitle("August 4") +  # Add the plot title
  ylim(0, 30) +  # Set y-axis limits to 0 and 30
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        strip.background = element_blank(),  # Remove background for facet labels
        strip.placement = "outside",  # Place facet labels outside the plot area
        panel.grid = element_blank())  # Remove the grid background

p2

```


###now make shannon's evenness plots

```{r}
library(dplyr)
library(tidyr)
library(vegan)

# Create an abundance table for each combination of Station and fraction
abund_table <- coverm_aug_mags %>%
  group_by(Station, fraction, host_genome) %>%
  summarise(total_abundance = sum(host_rel_ab), .groups = 'drop') %>%
  pivot_wider(names_from = host_genome, values_from = total_abundance, values_fill = list(total_abundance = 0))

# Function to calculate Shannon's Evenness for each row
calculate_shannon_evenness <- function(row) {
  abundance_values <- as.numeric(row[-c(1, 2)])  # Exclude the first two columns (Station and fraction)
  H <- diversity(abundance_values, index = "shannon")  # Shannon diversity index
  S <- sum(abundance_values > 0)  # Number of unique host genomes (non-zero abundance)
  E <- H / log(S)  # Shannon's Evenness
  return(E)
}

# Calculate Shannon's Evenness for each group
shannon_evenness_data <- abund_table %>%
  rowwise() %>%
  mutate(evenness = calculate_shannon_evenness(cur_data())) %>%
  select(Station, fraction, evenness) %>%
  ungroup()

```

```{r}
# Define the colors for each fraction
fraction_colors <- c("100" = "#29BF12", "3" = "#00A5CF", "53" = "#DE1A1A", "Whole" = "#FFBF00")

# Reorder the factors for Station and fraction
shannon_evenness_data$Station <- factor(shannon_evenness_data$Station, levels = c("WLE12", "WLE2", "WLE4"))
shannon_evenness_data$fraction <- factor(shannon_evenness_data$fraction, levels = c("Whole", "3", "53", "100"))

p2_evenness <- ggplot(shannon_evenness_data, aes(x = interaction(Station, fraction), y = evenness, fill = fraction)) +  # Set x to interaction of Station and fraction
  geom_boxplot(color = "black", width = 0.5) +  # Set the fill color by fraction and line color to black, adjust box width if needed
  geom_jitter(aes(fill = fraction), width = 0.2, alpha = 0.5, shape = 21, color = "black") +  # Add jittered points with fill color by fraction and black outline
  facet_grid(Station ~ fraction, scales = "free") +  # Use facet_grid to arrange panels as specified
  labs(x = "Station and Fraction", y = "Shannon Evenness") +
  scale_fill_manual(values = fraction_colors) +  # Use the named palette for coloring fractions
  theme_minimal() +
  ggtitle("Shannon Evenness - August 4") +  # Add the plot title
  ylim(0, 1) +  # Set y-axis limits to 0 and 1 (since evenness ranges from 0 to 1)
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        strip.background = element_blank(),  # Remove background for facet labels
        strip.placement = "outside",  # Place facet labels outside the plot area
        panel.grid = element_blank())  # Remove the grid background

p2_evenness

```

##sep host abundances
```{r}
setwd("~/Desktop/erie_story_r_work/sep_ma_mags_coverm_results/")
```

```{r}
files = list.files(pattern="*_ma_mags_relative_abundance.tsv",recursive=TRUE)

blast_hits <- function(x) {
  # Check if file is empty
  if (file.info(x)$size == 0) {
    message(paste0("Skipping empty file ", x))
    return(NULL)
  } else {
    tmp <- read.delim(x, header = FALSE, sep = "\t")
    tmp$sample <- gsub("", "", gsub("_ma_mags_relative_abundance.tsv", "", basename(x)))
    tmp$sample <- gsub("", "", gsub("sample_", "", basename(x)))
    return(tmp)
  }
}

tables <- Filter(NROW, lapply(files, blast_hits))
coverm_data <- rbindlist(tables)
```


##modify giant coverm file to contain meaningful columns
```{r}

names(coverm_data)[names(coverm_data) == 'V1'] <- "host_genome"
names(coverm_data)[names(coverm_data) == 'V2'] <- "host_rel_ab"
names(coverm_data)[names(coverm_data) == 'V3'] <- "host_mean_read_depth"
names(coverm_data)[names(coverm_data) == 'V4'] <- "host_covered_bases"
names(coverm_data)[names(coverm_data) == 'V5'] <- "host_variance"
names(coverm_data)[names(coverm_data) == 'V6'] <- "host_genome_length"
names(coverm_data)[names(coverm_data) == 'V7'] <- "host_rpkm"
names(coverm_data)[names(coverm_data) == 'V8'] <- "host_tpm"
names(coverm_data)[names(coverm_data) == 'V9'] <- "host_sample"

```

```{r}
coverm_data <- coverm_data[coverm_data$host_genome != 'unmapped',]
coverm_data <- coverm_data[coverm_data$host_genome != 'Genome',]
coverm_data$host_genome <- gsub("concatenated_concatenated_","",coverm_data$host_genome)
coverm_data$sample <- gsub("_ma_mags_relative_abundance.tsv","",coverm_data$sample)

coverm_data <- subset(coverm_data, grepl("contigs", host_genome))
```

##merge metadata and coverm data
```{r}
coverm_sep_mags <- merge(coverm_data, metadata, by = "sample")

coverm_sep_mags$host_genome <- gsub("_ma_contigs", "", coverm_sep_mags$host_genome)

coverm_sep_mags$host_rel_ab <- as.numeric(coverm_sep_mags$host_rel_ab)

```



```{r}
fraction_colors <- c("#29BF12", "#00A5CF", "#DE1A1A", "#FFBF00")

p2 <- ggplot(coverm_sep_mags, aes(x = factor(1), y = host_rel_ab)) +  # Set x to a constant to center the box plots
  geom_boxplot(aes(fill = fraction), color = "black", width = 0.5) +  # Set the fill color by fraction and line color to black, adjust box width if needed
  geom_jitter(aes(fill = fraction), width = 0.2, alpha = 0.5, shape = 21, color = "black") +  # Add jittered points with fill color by fraction and black outline
  facet_wrap(~ interaction(Station, fraction), scales = "free", strip.position = "bottom") +  # Use facet_wrap and place labels at the bottom
  labs(x = "Station and Fraction", y = "Relative Abundance") +
  scale_fill_manual(values = fraction_colors) +  # Use the defined palette for coloring fractions
  theme_minimal() +
  ggtitle("September 29") +  # Add the plot title
  ylim(0, 30) +  # Set y-axis limits to 0 and 30
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        strip.background = element_blank(),  # Remove background for facet labels
        strip.placement = "outside",  # Place facet labels outside the plot area
        panel.grid = element_blank())  # Remove the grid background

p2

```

```{r}
# Define the colors for each fraction
fraction_colors <- c("100" = "#29BF12", "3" = "#00A5CF", "53" = "#DE1A1A", "Whole" = "#FFBF00")

# Reorder the factors for Station and fraction
coverm_sep_mags$Station <- factor(coverm_sep_mags$Station, levels = c("WLE12", "WLE2", "WLE4"))
coverm_sep_mags$fraction <- factor(coverm_sep_mags$fraction, levels = c("Whole", "3", "53", "100"))

p2 <- ggplot(coverm_sep_mags, aes(x = factor(1), y = host_rel_ab)) +  # Set x to a constant to center the box plots
  geom_boxplot(aes(fill = fraction), color = "black", width = 0.5) +  # Set the fill color by fraction and line color to black, adjust box width if needed
  geom_jitter(aes(fill = fraction), width = 0.2, alpha = 0.5, shape = 21, color = "black") +  # Add jittered points with fill color by fraction and black outline
  facet_grid(Station ~ fraction, scales = "free") +  # Use facet_grid to arrange panels as specified
  labs(x = "Station and Fraction", y = "Relative Abundance") +
  scale_fill_manual(values = fraction_colors) +  # Use the named palette for coloring fractions
  theme_minimal() +
  ggtitle("September 29") +  # Add the plot title
  ylim(0, 30) +  # Set y-axis limits to 0 and 30
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        strip.background = element_blank(),  # Remove background for facet labels
        strip.placement = "outside",  # Place facet labels outside the plot area
        panel.grid = element_blank())  # Remove the grid background

p2

```


###now make shannon's evenness plots

```{r}
library(dplyr)
library(tidyr)
library(vegan)

# Create an abundance table for each combination of Station and fraction
abund_table <- coverm_sep_mags %>%
  group_by(Station, fraction, host_genome) %>%
  summarise(total_abundance = sum(host_rel_ab), .groups = 'drop') %>%
  pivot_wider(names_from = host_genome, values_from = total_abundance, values_fill = list(total_abundance = 0))

# Function to calculate Shannon's Evenness for each row
calculate_shannon_evenness <- function(row) {
  abundance_values <- as.numeric(row[-c(1, 2)])  # Exclude the first two columns (Station and fraction)
  H <- diversity(abundance_values, index = "shannon")  # Shannon diversity index
  S <- sum(abundance_values > 0)  # Number of unique host genomes (non-zero abundance)
  E <- H / log(S)  # Shannon's Evenness
  return(E)
}

# Calculate Shannon's Evenness for each group
shannon_evenness_data_sep <- abund_table %>%
  rowwise() %>%
  mutate(evenness = calculate_shannon_evenness(cur_data())) %>%
  select(Station, fraction, evenness) %>%
  ungroup()

```


##now get shannons evenness for two dates combined

```{r}
# Combine two dataframes using rbind in base R
combined_data <- rbind(coverm_aug_mags, coverm_sep_mags)

```

```{r}

# Create an abundance table for each combination of Station and fraction
abund_table <- combined_data %>%
  group_by(Station, fraction, host_genome) %>%
  summarise(total_abundance = sum(host_rel_ab), .groups = 'drop') %>%
  pivot_wider(names_from = host_genome, values_from = total_abundance, values_fill = list(total_abundance = 0))

# Function to calculate Shannon's Evenness for each row
calculate_shannon_evenness <- function(row) {
  abundance_values <- as.numeric(row[-c(1, 2)])  # Exclude the first two columns (Station and fraction)
  H <- diversity(abundance_values, index = "shannon")  # Shannon diversity index
  S <- sum(abundance_values > 0)  # Number of unique host genomes (non-zero abundance)
  E <- H / log(S)  # Shannon's Evenness
  return(E)
}

# Calculate Shannon's Evenness for each group
shannon_evenness_data_all <- abund_table %>%
  rowwise() %>%
  mutate(evenness = calculate_shannon_evenness(cur_data())) %>%
  select(Station, fraction, evenness) %>%
  ungroup()

```

```{r}

# Define the colors for each fraction
fraction_colors <- c("100" = "#29BF12", "3" = "#00A5CF", "53" = "#DE1A1A", "Whole" = "#FFBF00")

# Reorder the factors for Station and fraction
shannon_evenness_data_all$Station <- factor(shannon_evenness_data_all$Station, levels = c("WLE12", "WLE2", "WLE4"))
shannon_evenness_data_all$fraction <- factor(shannon_evenness_data_all$fraction, levels = c("Whole", "3", "53", "100"))

# Create the plot
p2_evenness <- ggplot(shannon_evenness_data_all, aes(x = interaction(Station, fraction), y = evenness, fill = fraction)) +  # Set x to interaction of Station and fraction
  geom_boxplot(color = "black", width = 0.5) +  # Set the fill color by fraction and line color to black, adjust box width if needed
  geom_jitter(aes(fill = fraction), width = 0.2, alpha = 0.5, shape = 21, color = "black") +  # Add jittered points with fill color by fraction and black outline
  facet_grid(Station ~ fraction, scales = "free") +  # Use facet_grid to arrange panels as specified
  labs(x = "Station and Fraction", y = "Shannon Evenness") +
  scale_fill_manual(values = fraction_colors) +  # Use the named palette for coloring fractions
  theme_minimal() +
  ggtitle("Shannon Evenness by Station and Fraction") +  # Add the plot title
  ylim(0, 1) +  # Set y-axis limits to 0 and 1 (since evenness ranges from 0 to 1)
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        strip.background = element_blank(),  # Remove background for facet labels
        strip.placement = "outside",  # Place facet labels outside the plot area
        panel.grid = element_blank())  # Remove the grid background

# Print the plot
print(p2_evenness)

```

```{r}

# Define the colors for each fraction
fraction_colors <- c("100" = "#29BF12", "3" = "#00A5CF", "53" = "#DE1A1A", "Whole" = "#FFBF00")

# Reorder the factor for fraction
shannon_evenness_data_all$fraction <- factor(shannon_evenness_data_all$fraction, levels = c("Whole", "3", "53", "100"))

# Create the plot
p2_evenness <- ggplot(shannon_evenness_data_all, aes(x = fraction, y = evenness, fill = fraction)) +  # Set x to fraction and y to evenness
  geom_boxplot(color = "black", width = 0.5) +  # Set the fill color by fraction and line color to black, adjust box width if needed
  geom_jitter(aes(fill = fraction), width = 0.05, alpha = 0.5, shape = 21, color = "black") +  # Add jittered points with fill color by fraction and black outline
  labs(x = "Sampling Fraction", y = "Shannon Evenness") +
  scale_fill_manual(values = fraction_colors) +  # Use the named palette for coloring fractions
  theme_minimal() +
  ggtitle("Shannon Evenness of Ma MAG diversity by Sampling Fraction") +  # Add the plot title
  ylim(0.7, 1) +  # Set y-axis limits to 0 and 1 (since evenness ranges from 0 to 1)
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        strip.background = element_blank(),  # Remove background for facet labels
        panel.grid = element_blank())  # Remove the grid background

# Print the plot
print(p2_evenness)

```

##Run tukey test on host diversity

```{r}
library(dplyr)

model <- lm( shannon_evenness_data_all$evenness ~ shannon_evenness_data_all$fraction )
ANOVA <- aov(model)
 
# Tukey test to study each pair of treatment :
TUKEY <- TukeyHSD(x=ANOVA, 'shannon_evenness_data_all$fraction', conf.level=0.95)

print(TUKEY)
```

```{r}
# Assuming shannon_evenness_data_all is your data frame

# Kruskal-Wallis Test
kruskal_test <- kruskal.test(evenness ~ fraction, data = shannon_evenness_data_all)
print(kruskal_test)

# Dunn's Test
install.packages("FSA")
library(FSA)
dunn_test <- dunnTest(evenness ~ fraction, data = shannon_evenness_data_all, method="bonferroni")
print(dunn_test)

# Welch's ANOVA
welch_anova <- oneway.test(evenness ~ fraction, data = shannon_evenness_data_all, var.equal = FALSE)
print(welch_anova)

# Pairwise t-tests
pairwise_t_test <- pairwise.t.test(shannon_evenness_data_all$evenness, shannon_evenness_data_all$fraction, p.adjust.method = "bonferroni")
print(pairwise_t_test)

# Linear Mixed-Effects Model (example with a random effect)
install.packages("lme4")
library(lme4)
# Assuming you have a column named 'random_effect' for the random effect
mixed_model <- lmer(evenness ~ fraction + (1|random_effect), data = shannon_evenness_data_all)
summary(mixed_model)

```

