---
title: "making_vp_abundance_plots_over_time_9_12_23"
author: "A.J Wing"
date: "2023-09-12"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_knit$set(root.dir = "~/Desktop/erie_story_r_work/first_paper_coverm_results/")

```

# Load necessary packages
```{r}
library(tidyr)
library(readr)
library(dplyr)
library(ggplot2)
library(viridis)
library(dendextend)
library(gridExtra)
library(vegan)
library(ape)
library(RColorBrewer)
library(jcolors)
library(showtext)
library(tidyverse)
library(data.table)
library(ggsignif)
```

###import metadata
```{r}
metadata <- read_tsv("~/Desktop/erie_story_r_work/2014_story_metadata_updated_8_9_22.tsv")
metadata$samples <- gsub("Sample_","",metadata$samples)
```



```{r}
files = list.files(pattern="*.tsv",recursive=TRUE)

coverm_results <- function(x) {
  # Check if file is empty
  if (file.info(x)$size == 0) {
    message(paste0("Skipping empty file ", x))
    return(NULL)
  } else {
    tmp <- read.delim(x, header = FALSE, sep = "\t")
    tmp$sample <- gsub("", "", gsub("_first_paper_vps_relative_abundance.tsv", "", basename(x)))
    return(tmp)
  }
}

tables <- Filter(NROW, lapply(files, coverm_results))
combined_blast_data <- rbindlist(tables)
```


##modify giant blast file to contain meaningful columns
```{r}

names(combined_blast_data)[names(combined_blast_data) == 'V1'] <- "genome"
names(combined_blast_data)[names(combined_blast_data) == 'V2'] <- "rel_ab"
names(combined_blast_data)[names(combined_blast_data) == 'V3'] <- "mean_coverage"
names(combined_blast_data)[names(combined_blast_data) == 'V4'] <- "covered_bases"
names(combined_blast_data)[names(combined_blast_data) == 'V5'] <- "variance"
names(combined_blast_data)[names(combined_blast_data) == 'V6'] <- "length"
names(combined_blast_data)[names(combined_blast_data) == 'V7'] <- "rpkm"
names(combined_blast_data)[names(combined_blast_data) == 'V8'] <- "tpm"
names(combined_blast_data)[names(combined_blast_data) == 'sample'] <- "samples"

```

##remove all unmapped rows and header rows
```{r}
combined_blast_data <- subset(combined_blast_data, !grepl("unmapped", genome))
combined_blast_data <- subset(combined_blast_data, !grepl("Genome", genome))
```

##keep columns of interest
```{r}
combined_blast_data <- combined_blast_data[, c("genome", "rel_ab", "mean_coverage", "covered_bases", "variance", "rpkm", "tpm", "samples")]
```



##import cluster file

```{r}
clusters <- read_tsv("~/Desktop/erie_story_r_work/final_parsed_clusters.tsv")
```

##modify cluster file so the vrhyme bins match the coverm outputs
```{r}
clusters$contig <- gsub("^vRhyme_(\\d+)__(\\d+)--(.*)$", "\\2_vRhyme_bin_\\1", clusters$contig)

names(clusters)[names(clusters) == 'contig'] <- "genome"
```

```{r}
virus_abundance_and_clusters <- merge(combined_blast_data, clusters, by ="genome", all=TRUE)
```

##combine with metadata 
```{r}
vp_abundance_clusters_and_metadata <- merge(virus_abundance_and_clusters, metadata, by ="samples", all=TRUE)
```



import viruses predicted to infect Ma
```{r}
aug_10kb_viruses_for_gephi <- read.csv("~/Desktop/erie_story_r_work/10kb_ma_only_aug4_infection_network_nodes_1_13_23.csv", header=TRUE)
```

```{r}
sep_10kb_viruses_for_gephi <- read.csv("~/Desktop/erie_story_r_work/10kb_ma_only_sep29_infection_network_nodes_1_13_23.csv", header=TRUE)
``` 

combine these
```{r}
all_10kb_ma_viruses <- rbind(aug_10kb_viruses_for_gephi, sep_10kb_viruses_for_gephi)

names(all_10kb_ma_viruses)[names(all_10kb_ma_viruses) == 'Name'] <- "genome"
```

#now narrow down to only predicted ma viruses
```{r}
vps_for_plotting <- merge(vp_abundance_clusters_and_metadata, all_10kb_ma_viruses, by ="genome", all=TRUE)
```

##keep only those that have values
```{r}
vps_for_plotting <- subset(vps_for_plotting, Type == "Virus")
```

```{r}

unique_count <- vps_for_plotting %>%
  summarise(unique_count = n_distinct(genome))

# Print the unique count
print(unique_count$unique_count)
```

```{r}
vps_for_plotting$cluster <- gsub("Cluster", "VP", vps_for_plotting$cluster)
```

```{r}

file_path <- "~/Desktop/erie_story_r_work/final_ma_vps_for_paper_1_9_22_23.tsv"


write.table(vps_for_plotting, file = file_path, sep = "\t", quote = FALSE, row.names = FALSE)
```

##only keep columns of interest for plotting
```{r}
vps_for_plotting_cleaned <- vps_for_plotting[,c("samples", "cluster", "genome", "length", "rel_ab", "Date", "Station.x", "fraction", "family")]

names(vps_for_plotting_cleaned)[names(vps_for_plotting_cleaned) == 'Station.x'] <- "Station"
```

```{r}
vps_for_plotting_cleaned$ordered_dates = factor(vps_for_plotting_cleaned$Date, levels=c("8-Jul-14","21-Jul-14","29-Jul-14","4-Aug-14","25-Aug-14","8-Sep-14","23-Sep-14","29-Sep-14","6-Oct-14","20-Oct-14","27-Oct-14"))
```

```{r}
# Create a new data frame containing rows where the "column_name" meets the condition
vps_for_plotting_cleaned_bloom_peaks <- subset(vps_for_plotting_cleaned, Date %in% c("4-Aug-14","29-Sep-14"))

vps_for_plotting_cleaned_bloom_peaks$rel_ab <- as.numeric(vps_for_plotting_cleaned_bloom_peaks$rel_ab)
```


```{r}
# Reorder the factor levels for Date so that 4-Aug-14 comes before 29-Sep-14
vps_for_plotting_cleaned_bloom_peaks$Date <- factor(vps_for_plotting_cleaned_bloom_peaks$Date, 
                                                        levels = c("4-Aug-14", "29-Sep-14"))

# Reorder the factor levels for fraction
vps_for_plotting_cleaned_bloom_peaks$fraction <- factor(vps_for_plotting_cleaned_bloom_peaks$fraction, 
                                                        levels = c("100", "53", "3", "Whole", "Viral"))

two_dates_taxonomy <- ggplot(vps_for_plotting_cleaned_bloom_peaks, aes(x = family, y = rel_ab, shape = Station, color = fraction)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_jitter(width = 0.2, height = 0.2, alpha = 0.7, size = 3) +
  scale_colour_jcolors(palette = "default") +
  facet_wrap(~ Date) +
  labs(x = "Viral Family", y = "Relative Abundance") +
  theme_bw() + 
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

two_dates_taxonomy

two_dates_taxonomy <- ggplot(vps_for_plotting_cleaned_bloom_peaks, 
                             aes(x = family, y = rel_ab, color = fraction)) +
  geom_point(size = 3, alpha = 0.7, 
             position = position_dodge(width = 0.7)) + # Add position_dodge here
  # Remove geom_jitter or apply jitter within dodge if needed
    scale_color_manual(values = c(
    "100" = "#29BF12",
    "3" = "#00A5CF",
    "53" = "#DE1A1A",
    "Viral" = "#574AE2",
    "Whole" = "#FFBF00"
  )) +
  facet_wrap(~ Date) +
  labs(x = "Viral Family", y = "Relative Abundance") +
  theme_bw() + 
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

two_dates_taxonomy

two_dates_taxonomy <- ggplot(vps_for_plotting_cleaned_bloom_peaks, 
                             aes(x = family, y = rel_ab, fill = fraction)) +
  geom_boxplot(position = position_dodge(width = 0.7), alpha = 0.7) +
  scale_fill_jcolors(palette = "default") +
  facet_wrap(~ Date) +
  labs(x = "Viral Family", y = "Relative Abundance") +
  theme_bw() + 
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

two_dates_taxonomy

```

```{r}
# Reorder the factor levels for Date so that 4-Aug-14 comes before 29-Sep-14
vps_for_plotting_cleaned_bloom_peaks$Date <- factor(vps_for_plotting_cleaned_bloom_peaks$Date, 
                                                        levels = c("4-Aug-14", "29-Sep-14"))

# Reorder the factor levels for fraction
vps_for_plotting_cleaned_bloom_peaks$fraction <- factor(vps_for_plotting_cleaned_bloom_peaks$fraction, 
                                                        levels = c("100", "53", "3", "Whole", "Viral"))

taxon_fraction_by_date <- ggplot(vps_for_plotting_cleaned_bloom_peaks, aes(x = family, y = rel_ab, shape = Station, color = fraction)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_jitter(width = 0.2, height = 0.2, alpha = 0.7, size = 3) +
  #geom_line(aes(color = fraction, group = cluster), size=0.75) +
  scale_colour_jcolors(palette = "default", breaks = c("100", "3", "53", "Viral", "Whole")) +
  facet_wrap(~ fraction + Date, nrow = 5, ncol = 2) +
  labs(x = "Viral Family", y = "Relative Abundance") +
  theme_bw() + 
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

taxon_fraction_by_date

```

```{r}
taxon_fraction_by_date <- ggplot(vps_for_plotting_cleaned_bloom_peaks, aes(x = family, y = rel_ab, shape = Station, color = fraction)) +
  geom_point(size = 3, alpha = 0.5) +
  geom_jitter(width = 0.2, height = 0.2, alpha = 0.5, size = 3) +
  scale_color_manual(values = c(
    "100" = "#29BF12",
    "3" = "#00A5CF",
    "53" = "#DE1A1A",
    "Viral" = "#574AE2",
    "Whole" = "#FFBF00"
  )) +
  facet_wrap(~ fraction + Date, nrow = 5, ncol = 2) +
  labs(x = "Viral Family", y = "Relative Abundance (% reads)") +
  theme_bw() + 
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

taxon_fraction_by_date
```

```{r}
vps_summary_by_date <- vps_for_plotting_cleaned_bloom_peaks %>%
  group_by(family, samples) %>%
  summarize(mean_rel_ab = mean(rel_ab, na.rm = TRUE)) %>%
  ungroup()

vps_summary_by_date <- vps_summary_by_date %>%
  left_join(
    vps_for_plotting_cleaned_bloom_peaks %>%
      select(samples, ordered_dates),
    by = "samples"
  )

vps_summary_by_date <- vps_summary_by_date %>%
  distinct()

mean_rel_ab_viral_family <- ggplot(vps_summary_by_date, aes(x = family, y = mean_rel_ab, color = ordered_dates)) +
  geom_boxplot(width = 0.8, alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.7) +
  scale_color_manual(name = "ordered_dates", values = c("4-Aug-14" = "#287D8EFF", "29-Sep-14" = "#287D8EFF")) +
  #scale_fill_manual(name = "ordered_dates", values = c("4-Aug-14" = "#287D8EFF", "29-Sep-14" = "#287D8EFF")) +
  facet_wrap(~ ordered_dates) +
  labs(x = "Viral Family", y = "Mean Rel. Ab. (% reads) of Viral Family per Sample") +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

mean_rel_ab_viral_family
```

```{r}

signif_mean_rel_ab_viral_family <- ggplot(vps_summary_by_date, aes(x = ordered_dates, y = mean_rel_ab, color = ordered_dates)) +
  geom_boxplot(width = 0.8, alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.7) +
  scale_color_manual(name = "ordered_dates", values = c("4-Aug-14" = "#287D8EFF", "29-Sep-14" = "#287D8EFF")) +
  #scale_fill_manual(name = "ordered_dates", values = c("4-Aug-14" = "#287D8EFF", "29-Sep-14" = "#287D8EFF")) +
  facet_wrap(~ family, nrow = 1) +
  labs(x = "Date", y = "Mean Rel. Ab. (% reads) of Viral Family per Sample") +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_signif(
    comparisons = list(c("4-Aug-14", "29-Sep-14")),
    map_signif_level = TRUE,
    test = "wilcox.test" # or "wilcox.test"
  )

signif_mean_rel_ab_viral_family

```

```{r}
# Reorder the factor levels for Date so that 4-Aug-14 comes before 29-Sep-14
vps_for_plotting_cleaned_bloom_peaks$Date <- factor(vps_for_plotting_cleaned_bloom_peaks$Date, 
                                                        levels = c("4-Aug-14", "29-Sep-14"))

taxon_fraction_by_date <- ggplot(vps_for_plotting_cleaned_bloom_peaks, aes(x = Date, y = rel_ab, color = family)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_jitter(width = 0.2, height = 0.2, alpha = 0.7, size = 3) +
  geom_line(aes(color = family, group = cluster), size=0.75) +
  #scale_colour_jcolors(palette = "default") +
  facet_wrap(~ Station + fraction, nrow = 3, ncol = 5) +
  labs(x = "Viral Family", y = "Relative Abundance") +
  theme_bw() + 
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

taxon_fraction_by_date

```


```{r}
ggplot(vps_for_plotting_cleaned_bloom_peaks, aes(x = family, y = rel_ab)) +
  geom_point(size = 3, shape = 21, alpha = 0.8, fill = '#287D8EFF', color = '#287D8EFF') +
  geom_jitter(width = 0.1, height = 0.1, alpha = 0.8, size = 3, shape = 21, 
              fill = '#287D8EFF', color = '#287D8EFF') +
  facet_wrap(~ Date + fraction + Station) +
  labs(x = "Viral Family", y = "Relative Abundance") +
  theme_bw() + 
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )


```

```{r}
ggplot(vps_for_plotting_cleaned_bloom_peaks, aes(x = family, y = rel_ab, shape = Station)) +
  geom_point(size = 3, alpha = 0.8, fill = '#287D8EFF', color = '#287D8EFF') +
  geom_jitter(width = 0.1, height = 0.1, alpha = 0.8, size = 3, shape = 21, 
              fill = '#287D8EFF', color = '#287D8EFF') +
  facet_grid(fraction ~ Date) +
  labs(x = "Viral Family", y = "Relative Abundance") +
  theme_bw() + 
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )


```








```{r}
# Create a new data frame containing rows where the "column_name" meets the condition
vps_for_plotting_cleaned_subset_1 <- subset(vps_for_plotting_cleaned, cluster %in% c("VP_1884","VP_2860","VP_2719", "VP_1873", "VP_2601", "VP_583", "VP_2852", "VP_885", "VP_1929", "VP_2297"))

```

##make sure ranges plotted are numeric
```{r}
vps_for_plotting_cleaned_subset_1$rel_ab <- as.numeric(vps_for_plotting_cleaned_subset_1$rel_ab)
```

```{r}
vps_for_plotting_subset_1_plot <- ggplot(vps_for_plotting_cleaned_subset_1, aes(x=ordered_dates, y=rel_ab, group=cluster, color=cluster)) +
 geom_point(aes(color = cluster), alpha=0.8, size=6) +
 geom_line(aes(color = cluster, group = cluster), size=0.75) +
  scale_color_manual(values = c('#20DE8B', '#CCDE8B', '#FFDE8B', '#FFA88B', '#FF6A8B', '#FF6AD5', '#C874AA', '#AD8CFF', '#966BFF', '#90CFFF')) +
  ggtitle("LE Microcystis Viral Populations with highest relative abundance") +
  xlab("Sampling Date") +
  ylab("TPM Relative Abundance (%)") +
  guides(colour = guide_legend(override.aes = list(size=6))) +
  theme(axis.text.x = element_text(angle =45, hjust = 1)) +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"), 
        panel.border = element_rect(colour = "black", fill=NA, size=1), 
        axis.title=element_text(size=14),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        plot.title = element_text(size = 16, hjust = 0.5, colour = "black"),
        axis.text.y = element_text(size=12, colour = "black"),
        axis.text.x  = element_text(size=12, colour="black"),
        plot.margin = unit(c(1, 1, 1, 1),"lines"))
        plot.margin = margin(t = 1,  # Top margin
                             r = 1,  # Right margin
                             b = 3,  # Bottom margin
                             l = 2,  # Left margin
                             unit = "cm")


vps_for_plotting_subset_1_plot
faceted_plot_1 <- vps_for_plotting_subset_1_plot + facet_grid(fraction ~ Station)
faceted_plot_1
```

```{r}
png("~/Desktop/erie_story_r_work/most_abundant_ma_vps_through_time_and_space_9_14_23.png", width=1000, height=600)
faceted_plot_1
dev.off()
```



```{r}
pdf("~/Desktop/erie_story_r_work/most_abundant_ma_vps_through_time_and_space_9_14_23.pdf", width=9, height=6)
faceted_plot_1
dev.off()
```



```{r}
vps_for_plotting_cleaned <- vps_for_plotting_cleaned %>%
  group_by(cluster) %>%
  mutate(number_dates = n_distinct(Date))
```

```{r}
vps_for_plotting_subset_1_plot <- ggplot(vps_for_plotting_cleaned_subset_1, aes(x=ordered_dates, y=rel_ab, group=cluster, color=cluster)) +
 geom_point(aes(color = cluster), alpha=0.8, size=6) +
 geom_line(aes(color = cluster, group = cluster), size=0.75) +
  scale_color_manual(values = c('#20DE8B', '#CCDE8B', '#FFDE8B', '#FFA88B', '#FF6A8B', '#FF6AD5', '#C874AA', '#AD8CFF', '#966BFF', '#90CFFF')) +
  ggtitle("LE Microcystis Viral Populations with highest relative abundance") +
  xlab("Sampling Date") +
  ylab("TPM Relative Abundance (%)") +
  guides(colour = guide_legend(override.aes = list(size=6))) +
  theme(axis.text.x = element_text(angle =45, hjust = 1)) +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"), 
        panel.border = element_rect(colour = "black", fill=NA, size=1), 
        axis.title=element_text(size=14),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        plot.title = element_text(size = 16, hjust = 0.5, colour = "black"),
        axis.text.y = element_text(size=12, colour = "black"),
        axis.text.x  = element_text(size=12, colour="black"),
        plot.margin = unit(c(1, 1, 1, 1),"lines"))
        plot.margin = margin(t = 1,  # Top margin
                             r = 1,  # Right margin
                             b = 3,  # Bottom margin
                             l = 2,  # Left margin
                             unit = "cm")


vps_for_plotting_subset_1_plot
faceted_plot_1 <- vps_for_plotting_subset_1_plot + facet_wrap(vars(fraction))
faceted_plot_1
```




##Try grouping by station in addition to VP

```{r}


vps_for_plotting_subset_1_plot <- ggplot(vps_for_plotting_cleaned_subset_1, aes(x=ordered_dates, y=rel_ab, group=interaction(cluster, Station), color=cluster, shape=Station)) +
  geom_point(aes(color = cluster), alpha=0.6, size=4) +
  geom_line(aes(color = cluster, group = interaction(cluster, Station)), alpha=0.6, size=0.75) +
  scale_color_manual(values = c('#20DE8B', '#CCDE8B', '#FFDE8B', '#FFA88B', '#FF6A8B', '#FF6AD5', '#C874AA', '#AD8CFF', '#966BFF', '#90CFFF')) +
  scale_shape_manual(values = c(15, 16, 17)) +  
  ggtitle("LE Microcystis VPs Top Ten Relative Abundance") +
  xlab("Sampling Date") +
  ylab("Relative Abundance (%)") +
  guides(colour = guide_legend(override.aes = list(size=6))) +
  theme(axis.text.x = element_text(angle =45, hjust = 1)) +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"), 
        panel.border = element_rect(colour = "black", fill=NA, size=1),
        strip.text = element_text(size = 12),
        strip.background = element_rect(colour= "black", fill = "lemon chiffon", linetype= "solid"),
        axis.title=element_text(size=14),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        plot.title = element_text(size = 16, hjust = 0.5, colour = "black"),
        axis.text.y = element_text(size=12, colour = "black"),
        axis.text.x  = element_text(size=12, colour="black"),
        plot.margin = unit(c(1, 1, 1, 1),"lines"))
        plot.margin = margin(t = 1,  # Top margin
                             r = 1,  # Right margin
                             b = 3,  # Bottom margin
                             l = 2,  # Left margin
                             unit = "cm")

vps_for_plotting_subset_1_plot
faceted_plot_1 <- vps_for_plotting_subset_1_plot + facet_wrap(vars(fraction))
faceted_plot_1

```

```{r}
png("~/Desktop/erie_story_r_work/most_abundant_ma_vps_through_time_and_space_9_18_23.png", width=1000, height=600)
faceted_plot_1
dev.off()
```



```{r}
pdf("~/Desktop/erie_story_r_work/most_abundant_ma_vps_through_time_and_space_9_18_23.pdf", width=9, height=6)
faceted_plot_1
dev.off()
```